An Executable Formal Semantics For A Functional Actor Language

### 2017-12-21 paper/paper.tex

Conditional reduction rules can also be easily specified.
For example, the branching rule (br) is specified as two K rules
\begin{verbatim}
  rule br(nil, V1:Val, V2:Val) => V2
  rule <k> br(V:Val, V1:Val, V2:Val) => V1 ... </k>
       <definedaddr> X </definedaddr>
    requires notBool((isId(V) andBool notBool(V in X))
                              orBool V ==K nil)
\end{verbatim}
The first rule says if the first argument is \texttt{nil}, the branching 
operator \texttt{br} returns its third argument.
The second rule is a bit complicated.
It says if the first argument is neither \texttt{nil} nor 
unknown addresses, the branching operator \texttt{br} returns its second 
argument.
The \texttt{requires} keyword is used to specify rewriting conditions that 
decide when a rewrite rule can be applied.

The single-step transition relation $\mapsto$ on actor configurations is 
similarly specified using K rules.
For example, the \texttt{<init: a, a1>} transition rule is specified as
\small
\begin{verbatim}
  rule <actor>
         <k> initbeh(A1, V:Val) => nil ... </k>
         <id> A </id>
       </actor>
       <actor>
         <k> uninit(A) => ready(V) ... </k>
         <id> A1 </id>
       </actor>
\end{verbatim}
\normalsize
The rule involves two actors as there are two actor cells 
\texttt{<actor>}\ \dots\ \texttt{</actor>} in the rule.
The first actor is the \emph{initializer} actor while the second is the actor 
\emph{being initialized}.
K supports to only specify components that are relevant to the rule. This is 
known as the \emph{framing principle}.
The initializer actor executes \texttt{initbeh(A1, V:Val)} and gets value 
\texttt{nil}, as specified in the rule.
The rest computation or so-called \emph{reduction context} is unchanged and 
kept in 
the following~``\texttt{...}''.

Reduction context can be mentioned explicitly when needed.
For example, in the following \texttt{<bec: a, a1>} rule,
\small
\begin{verbatim}
  rule <actor>
         <k> (become(V:Val) ~> RestComputation) => ready(V) </k>
         ...
       </actor>
       (.Bag => <actor>
                  <k> nil ~> RestComputation </k>
                  <id> !A1:Id </id>
                </actor>)
       <definedaddr> Ads => Ads ListItem(!A1) </definedaddr>
\end{verbatim}
\normalsize
the reduction context \texttt{RestComputation} is explicitly passed to a 
newly-generated actor with a fresh identity \verb|!A1|.
The tilde arrow expression \verb|T ~> R| 
means the decomposition \verb|R[T]| with reduction context \texttt{R} and redex 
\texttt{T}.

>>>>

Conditional reduction rules can also be easily specified.
For example, the branching rule (br) is specified as two K rules
\begin{verbatim}
  rule br(nil, V1:Val, V2:Val) => V2
  rule <k> br(V:Val, V1:Val, V2:Val) => V1 ... </k>
       <definedaddr> X </definedaddr>
    requires notBool((isId(V) andBool notBool(V in X))
                              orBool V ==K nil)
\end{verbatim}
The first rule says if the first argument is \texttt{nil}, the branching 
operator \texttt{br} returns its third argument.
The second rule is a bit complicated.
It says if the first argument is neither \texttt{nil} nor an
unknown address, the branching operator \texttt{br} returns its second
argument.
The \texttt{requires} keyword is used to specify rewriting conditions that 
decide when a rewrite rule can be applied.

The single-step transition relation $\mapsto$ on actor configurations is 
similarly specified using K rules.
For example, the \texttt{<init: a, a1>} transition rule is specified as
\small
\begin{verbatim}
  rule <actor>
         <k> initbeh(A1, V:Val) => nil ... </k>
         <id> A </id>
       </actor>
       <actor>
         <k> uninit(A) => ready(V) ... </k>
         <id> A1 </id>
       </actor>
\end{verbatim}
\normalsize
The rule involves two actors as there are two actor cells 
\texttt{<actor>}\ \dots\ \texttt{</actor>} in the rule.
The first actor is the \emph{initializer} actor while the second is the actor 
\emph{being initialized}.
K supports to only specify components that are relevant to the rule. This is 
known as the \emph{framing principle}.
The initializer actor executes \texttt{initbeh(A1, V:Val)} and gets value 
\texttt{nil}, as specified in the rule.
The rest of the computation or so-called \emph{reduction context} is unchanged and
kept in 
the following~``\texttt{...}''.

Reduction contexts can be mentioned explicitly when needed.
For example, in the following \texttt{<bec: a, a1>} rule,
\small
\begin{verbatim}
  rule <actor>
         <k> (become(V:Val) ~> RestComputation) => ready(V) </k>
         ...
       </actor>
       (.Bag => <actor>
                  <k> nil ~> RestComputation </k>
                  <id> !A1:Id </id>
                </actor>)
       <definedaddr> Ads => Ads ListItem(!A1) </definedaddr>
\end{verbatim}
\normalsize
the reduction context \texttt{RestComputation} is explicitly passed to a 
newly-generated actor with a fresh identity \verb|!A1|.
The tilde arrow expression \verb|T ~> R| 
means the decomposition \verb|R[T]| with reduction context \texttt{R} and redex 
\texttt{T}.
